On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-08-17 manuscripts/PNAS/chazthm_pnas.tex
chaz

@@

@@

@@

%We have demonstrated the uniqueness and stability of solutions to Prob. \ref{InverseProblem} and the related Prob. \ref{OptimizationProblem}, 
To this end, we have conclusively verified that inference in this parsimonious model of (noisy) natural data is a well-posed problem in the sense of Hadamard \cite{Hadamard1902}, a long-standing necessary property of reductionist models in science. This notion summarizes the idea that inferences from observations should be robust to the inevitable uncertainty in measurement. In this regard, we demonstrate rigorously that, given sufficient data and suitable dictionary, Prob.~\ref{OptimizationProblem} reduces to an instance of Prob.~\ref{InverseProblem} to which applies our main result (Thm.~\ref{DeterministicUniquenessTheorem}): every dictionary and sparse codes solving Prob.~\ref{InverseProblem} are equivalent up to inherent relabeling/scaling ambiguities and a discrepancy (error) that scales linearly with the measurement noise or modeling inaccuracy; and we provide explicit constants expressed in terms of these latent parameters.
%In other words, a small perturbation of the data should result in only slightly different inferred parameters. 
%Moreover, we show that even if the meta-parameter for the number of dictionary elements is overestimated, a subset of parameters may still be identifiable up to noise. 
%This work justifies the application of the sparse linear coding model to blind source separation problems, wherein the goal is to infer the generating dictionary and sparse codes from noisy measurements, as in \eqref{LinearModel}. We also collect a set of useful mathematical tools and basic facts for future research.

@@

 [*** TODO: PUT in paper sprinkled TIGHTNESS and COMPUTABILITY ***]

>>>>

Before moving on to our proofs, we discuss briefly the computability and tightness of our results. [* DO THIS *]
%\pagebreak

@@

[* PUT COMPUTABILITY IN HERE *]

@@

We remark that if $k$ is held fixed, or if the size of the support set of reconstructing codes is known to be polynomial in $m$ and $k$, then a practical (polynomial) amount of data is enough to determine the model.\footnote{In the latter case, the pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} requires only a polynomial number of samples distributed over a polynomial number of supports.}

@@

%We have demonstrated the uniqueness and stability of solutions to Prob. \ref{InverseProblem} and the related Prob. \ref{OptimizationProblem}, 
To this end, we have conclusively verified that inference in this parsimonious model of (noisy) natural data is a well-posed problem in the sense of Hadamard \cite{Hadamard1902}, a long-standing necessary property of reductionist models in science. This notion summarizes the idea that inferences from observations should be robust to the inevitable uncertainty in measurement. In this regard, we demonstrate rigorously that, given sufficient data and suitable dictionary, Prob.~\ref{OptimizationProblem} reduces to an instance of Prob.~\ref{InverseProblem} to which applies our main result (Thm.~\ref{DeterministicUniquenessTheorem}): every dictionary and sparse codes solving Prob.~\ref{InverseProblem} are equivalent up to inherent relabeling/scaling ambiguities and a discrepancy (error) that scales linearly with the measurement noise or modeling inaccuracy; and we provide explicit, computable constants expressed in terms of these latent parameters.
%In other words, a small perturbation of the data should result in only slightly different inferred parameters. 
%Moreover, we show that even if the meta-parameter for the number of dictionary elements is overestimated, a subset of parameters may still be identifiable up to noise. 
%This work justifies the application of the sparse linear coding model to blind source separation problems, wherein the goal is to infer the generating dictionary and sparse codes from noisy measurements, as in \eqref{LinearModel}. We also collect a set of useful mathematical tools and basic facts for future research.

@@
